\TheTitle

### 2019-09-20 Paper1/results.tex
smaller changes results

Recall that the numerical flux \eqref{eq:numericalFluxIPM} uses a quadrature rule to approximate integrals. We start by investigating the effects this quadrature has on the solution accuracy. For this, we run the IPM method with a moment order ranging from three to seven using a Clenshaw-Curtis quadrature rule with level three (i.e. 9 quadrature points) and level four (i.e. 17 quadrature points). A comparison of the error obtained with these two quadrature levels is given in Figure~\ref{fig:ErrorDifferentQuad}. 
\begin{itemize}
\item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate. This behavior results from aliasing effects in the numerical flux, which dominate the accuracy level.
\item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy. Note that this holds for the expectation value until a truncation order of $N=5$ and for the variance for $N=4$. Hence, the variance is more sensitive to aliasing errors.
\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, whereas the error of the variance in Figure~\ref{fig:ErrorDifferentQuadB} shows that IPM with 9 quadrature points yields an improved variance results over SC$^9$ even with only four moments. Hence, the number of moments needed for IPM to obtain a certain variance error is significantly smaller than the number of quadrature points needed for SC. This result can be observed throughout the results of this work. Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly. However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
\end{itemize}
Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order 9. Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well. To account for the fact that sparse grids need to be used in high dimensional problems, we use Clenshaw-Curtis quadrature rules to compute the different solutions. Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points. All adaptive methods use gPC polynomials of order 2 to 9 (i.e. 3 to 10 moments). Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule. The remaining methods have been computed with a level 4 quadrature rule.

@@

First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration. The standard SG method however preserves positivity of mass, energy and pressure. The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}. When comparing intrusive methods without acceleration techniques as well as SC, the following properties stick out:
\begin{itemize}
\item SG comes at a heavily reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
\item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns. Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
\end{itemize}
The proposed acceleration techniques show the following behavior:
\begin{itemize}
\item The one shot IPM (osIPM) method proposed in in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the one-shot idea, the method is denoted by \textit{adaptive one-shot IPM} (adosIPM). This method reaches the steady state IPM solution with a faster runtime than SG.
\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by retardation adosIPM (readosIPM), which further decreases runtime. Here, we use two different strategies: First, we steadily increase the maximal truncation order when the residual approaches zero. To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order. The residual values are then determined to be $0.00006,0.00003,0.000022$ and $0.00002$. The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$. Strategy 1 is depicted in red, Strategy 2 is depicted in green. It can be seen that both approaches reach the IPM$_9$ error for the same run time. Hence, we deduce that a naive choice of the refinement retardation strategy suffices to yield a satisfactory behavior.
\end{itemize}
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_ERho}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_ERho}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_ERho}.png}
		\label{fig:sub1}
	\end{subfigure}

		\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_VarRho}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_VarRho}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_VarRho}.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
	\label{fig:ERhoVarRho5}
\end{figure}
Let us finally take a look at the expectation value and variance computed with different methods. All results are depicted for a zoomed view around the airfoil. Figure~\ref{fig:ERhoVarRho5} shows the expectation value (first row) and variance (second row) computed with $5$ quadrature points for SC and $5$ moments for SG and IPM. One can observe the following
\begin{itemize}
\item All methods yield non-physical step-like profiles of the expectation value and variance along the airfoil. 
\item The jump position of the intrusive solution profiles capture the exact behavior more accurately.
\end{itemize}
The readosIPM results as well as the corresponding refinement levels are depicted in Figure~\ref{fig:readosIPMEVar}.
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_ERho}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_VarRho}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_refinementLevel}.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{2-9}$}
	\label{fig:readosIPMEVar}
\end{figure}

>>>>

Recall that the numerical flux \eqref{eq:numericalFluxIPM} uses a quadrature rule to approximate integrals. We start by investigating the effects this quadrature has on the solution accuracy. For this, we run the IPM method with a moment order ranging from three to seven using a Clenshaw-Curtis quadrature rule with level three (i.e. 9 quadrature points) and level four (i.e. 17 quadrature points). A comparison of the error obtained with these two quadrature levels is given in Figure~\ref{fig:ErrorDifferentQuad}. To denote the quadrature level, we use a superscript and the moment order is denoted by a subscript.
\begin{itemize}
\item When for example comparing the error obtained with IPM$_8^3$ and IPM$_8^4$ it can be seen that the error stagnates when the chosen quadrature is not sufficiently accurate. This behavior results from aliasing effects in the numerical flux, which dominate the accuracy level.
\item If the truncation order is sufficiently small, both quadrature levels yield the same accuracy. Note that this holds for the expectation value until a truncation order of $N=5$ and for the variance for $N=4$. Hence, the variance is more sensitive to aliasing errors.
\item Figure~\ref{fig:ErrorDifferentQuadA} reveals that the IPM error of E[$\rho$] with 9 quadrature points stagnates at the error level of SC$^9$, i.e. the aliasing error heavily affects the accuracy. This behavior becomes more dominant when looking at the variance error in Figure~\ref{fig:ErrorDifferentQuadB}. Here, IPM$_8^4$ yields a significantly improved result compared to IPM$_8^3$. Again, the IPM$_8^3$ result stagnates at the SC$^9$ accuracy level, which can be reached with IPM when only using four moments (combined with a sufficiently accurate quadrature level). Hence, the number of moments needed for IPM to obtain a certain variance error is significantly smaller than the number of quadrature points needed for SC. This result can be observed throughout the results of this work. Especially for high dimensional problems, this potentially decreases the number of unknowns to reach a certain accuracy level significantly. However note, that since the numerical flux evaluation requires $O(N\cdot Q)$ operations, the numerical costs will depend on the chosen quadrature rule.
\end{itemize}
Let us now compare stochastic-Collocation with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order 9. Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well. To account for the fact that sparse grids need to be used in high dimensional problems, we use Clenshaw-Curtis quadrature rules to compute the different solutions. Stochastic-Collocation uses quadrature levels $2,3$ and $4$, which corresponds to $5,9$ and $17$ quadrature points. All adaptive methods use gPC polynomials of order 2 to 9 (i.e. 3 to 10 moments). Order 2 uses a level 2, orders 3 to 6 use a level 3 and orders 8 and 9 use a level 4 quadrature rule. When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the truncation order is decreased, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased. The remaining methods have been computed with a level 4 quadrature rule.

@@

First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration. The standard SG method however preserves positivity of mass, energy and pressure. The change of the relative L$^2$ error during the iteration to the steady state has been recorded in Figure~\ref{fig:L2ErrorSolution}. When comparing intrusive methods without acceleration techniques as well as SC, the following properties stick out:
\begin{itemize}
\item Compared to IPM, stochastic-Galerkin comes at a heavily reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
\item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns. Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
\end{itemize}
The proposed acceleration techniques show the following behavior:
\begin{itemize}
\item The one shot IPM (osIPM) method proposed in in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the one-shot idea, the method is denoted by \textit{adaptive one-shot IPM} (adosIPM). This method reaches the steady state IPM solution with a faster runtime than SG.
\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by retardation adosIPM (readosIPM), which further decreases runtime. Here, we use two different strategies: First, we steadily increase the maximal truncation order when the residual approaches zero. To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order. The residual values are then determined to be $0.00006,0.00003,0.000022$ and $0.00002$. The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$. Strategy 1 is depicted in red, Strategy 2 is depicted in green. It can be seen that both approaches reach the IPM$_9$ error for the same run time. Hence, we deduce that a naive choice of the refinement retardation strategy suffices to yield a satisfactory behavior.
\end{itemize}
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_ERho}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_ERho}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_ERho}.png}
		\label{fig:sub1}
	\end{subfigure}

		\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{EulerSC5-2Res1e-6_VarRho}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{sg_4_9_VarRho}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/{osIPM4-4_VarRho}.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_5$, IPM$_5$ (from left to right).}
	\label{fig:ERhoVarRho5}
\end{figure}
Let us finally take a look at the expectation value and variance computed with different methods. All results are depicted for a zoomed view around the airfoil. Figure~\ref{fig:ERhoVarRho5} shows the expectation value (first row) and variance (second row) computed with $5$ quadrature points for SC and $5$ moments for SG and IPM. One can observe the following
\begin{itemize}
\item All methods yield non-physical step-like profiles of the expectation value and variance along the airfoil. 
\item The jump position of the intrusive solution profiles capture the exact behavior more accurately.
\end{itemize}
The readosIPM results as well as the corresponding refinement levels are depicted in Figure~\ref{fig:readosIPMEVar}. One observes that the solution no longer shows the previously observed discontinuous profile and yield a satisfactory agreement with the reference solution depicted in Figure~\ref{fig:referenceSolution}. The refinement level shows that away from the airfoil, a refinement level of 0 (i.e. a truncation order of 2) suffices to yield the IPM$_9$ solution. The region with a high variance requires a refinement level of 7 (truncation order 9).
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_ERho}.png}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_VarRho}.png}
		\label{fig:sub2}
	\end{subfigure}%
	\begin{subfigure}{0.33\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots10/{euler2D_nacaCoarse_refadosipm_n2-9_sg2-4_s05_aoa_oneRet_refinementLevel}.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$, Var$[\rho]$ and refinement level for readosIPM$_{2-9}$}
	\label{fig:readosIPMEVar}
\end{figure}
